Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-10-22 refreport/response.tex

\documentclass[11pt]{article}

\usepackage{fullpage}
\usepackage{natbib}
\usepackage{hyperref, breakurl, color}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\newcommand{\blue}[1]{\textcolor{blue}{#1}}

\newcommand{\code}[1]{\textsf{#1}}
\newcommand{\enzo}{\code{Enzo}}

\def\apj{ApJ}
\def\apjl{ApJL}
\def\apjs{ApJS}
\def\mnras{MNRAS}
\def\aj{Ast.J.}
\def\nat{Nature}
\def\na{New.A.}
\newcommand{\aap}{A\&A}

\begin{document}

\begin{center} 
\bfseries{
\begin{large}
Response to referee report for manuscript  ApJ/ApJS92163
\end{large}
}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This is the response to the referee's report for ApJ/ApJS manuscript
\#92163, ``Enzo: An Adaptive Mesh Refinement Code for Astrophysics,''
by the Enzo collaboration (Bryan et al.).  We thank the referee for
their feedback, and address each comment separately below.

\section{Test problems}

\red{(Add response regarding test problems here)}

\section{Paper length and breakdown}

Finally, we would like to address the referee's feedback regarding the
paper's length and current organization.  The referee's comment is:

\begin{verbatim} 
Given the paper's length as it currently stands, it seems appropriate
to split the paper into two. One paper should document the basic
algorithms (e.g. AMR, hydro, MHD and gravity implementations) along
with an appropriate range of tests based on standard tests from the
literature. The second paper should document the microphysics options
(chemistry, radiation, thermal conductivity, etc.).
\end{verbatim}

We respectfully disagree with the referee regarding this point --
splitting the paper in two does not make sense, will not shorten the
overall page count, and will not improve the clarity of presentation. One of the strengths
of the \enzo\ code is the breadth of physics packages that are
available in the publicly-available version -- unlike, for example,
Gadget and Gadget-2 \citep{2001NewA....6...79S, 2005MNRAS.364.1105S},
where the only physics packages that are publicly available are dark
matter and adiabatic hydrodynamics.  Given this, it is logical that
the totality of the code's capabilities be documented in a single
place, for reference by future generations of computational scientists.

Regarding the paper's length, we argue that it is not excessively long
given the topics it covers, and thus does not need to be split into
multiple parts.  At present, the \enzo\ method paper is \red{60 pages
in ApJ format (BWO: we will need to update this before submission)}.
For comparison:

\begin{itemize}
\item The paper for the open-source FLASH code
\citep{2000ApJS..131..273F} is 62 pages long -- almost identical in
length to the \enzo\ method paper.  At the time of the FLASH method paper's
publication, it covered the full feature set of the FLASH code, which
is comparable to the breadth of topcs covered in the \enzo\ method
paper.

\item The AREPO method paper \citep{2010MNRAS.401..791S} is 60 pages
long, covers a narrower range of topics, and the code is
closed-source.

\item The paper for the open-source Athena code
\citep{2008ApJS..178..137S} is 41 pages long, but only documents the
hydrodynamics and MHD algorithms.  At the time of the code's release,
this paper documented the entire feature set of the publicly-available
code.

\item The paper for the open-source ZEUS-MP 2 code
\citep{2006ApJS..165..188H} is 41 pages long, but again covers a
substantially narrower range of physics modules.  The entirety of the
publicly-available code is documented in this paper.
\end{itemize}

The only \textit{comparable} open-source code that we are aware of
whose method paper is broken up into multiple segments is the series
of three ZEUS-2D papers \citep{1992ApJS...80..753S,
1992ApJS...80..791S, 1992ApJS...80..819S}.  We argue, however, that
this is not a useful comparison, as the three papers were published in
the same issue of the Astrophysical Journal in 1992, and were
published back-to-pack as consecutive articles in the same volume.  In
a time when the primary mode of consumption of articles was via paper
journals, this is effectively a single paper (with a length of
approximately 85 pages, no less).

We further note that splitting the paper up into two separate pieces
provides structural challenges, and will serve to increase the total
page count.  Structurally, Section 2 (the ``overview'' section) would
need to be split into two pieces, and the component that discusses the
microphysics will be somewhat confusing on its own, as these modules
act as sink and source terms for the fluid equations that would in
principle be described in a separate paper.  It's not clear where
Section 10 (``Analysis''), Section 12 (``Parallel strategy and
performance''), and Section 13 (``Code development methodology'')
would logically go -- some of these sections apply to all physics
packages in the code (e.g., Section 12), and others to none in
particular (e.g., Section 13).  Similarly, Section 11.1 (``Verifying
and validating the Enzo code'') describes the \enzo\ test suite, which
is used for all aspects of code testing (including both the
fluid/gravity/AMR algorithms \textit{and} the microphysics), and also
for code comparisons that test many different aspects of the code,
often in combination.  In order for these sections to make sense, and
to address test problems that are split between two papers, several of
these sections would need to be either duplicated or require an
extensive amount of cross-referencing between papers Furthermore,
separate introductions, discussions, and conclusions will by necessity
be mutually referential and, in some sense, duplicates of each other.
Together, this will not serve to either decrease page count or
increase clarity for the reader.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage

\bibliographystyle{apj}
\bibliography{response}

\end{document}

### 2013-10-22 acknowledgments.tex

This work has been supported by the National Science Foundation by
grants
AAG-0808184 (DRR),
AAG-1109008 (DRR),
ACI-9619019 (MLN),
ASC-9313135 (MLN),
AST-9803137 (MLN), 
AST-0307690 (MLN), 
AST-0407176 (RC),
AST-0407368 (SS, EJH),
AST-0507521 (RC), 
AST-0507717 (MLN), 
AST-0507768 (AK),
AST-0529734 (TA),
AST-0607675 (AK),
AST-0702923 (EJH),
AST-0707474 (BDS), 
AST-0708960 (MLN), 
AST-0807075 (TA),
AST-0807215 (JB),
AST-0808184 (MLN, AK),
AST-0808398 (TA),
AST-0908740 (AK, DCC),
AST-0908819 (BWO), 
AST-0955300 (NJG),
AST-1008134 (GB), 
AST-1009802 (JSO), 
AST-1102943 (MLN),
AST-1106437 (JB),
AST-1210890 (GB),
AST-1211626 (JHW),
OCI-0832662 (BWO, MLN),
OCI-0941373 (BWO),
PHY-1104819 (MLN, JB),
the CI TraCS fellowship (OCI-1048505; MJT),
and the Graduate Research Fellowship program (NJG; SWS).

>>>>

This work has been supported by the National Science Foundation by
grants
AAG-0808184 (DRR),
AAG-1109008 (DRR),
ACI-9619019 (MLN),
ASC-9313135 (MLN),
AST-9803137 (MLN), 
AST-0307690 (MLN), 
AST-0407176 (RC),
AST-0407368 (SS, EJH),
AST-0507521 (RC), 
AST-0507717 (MLN), 
AST-0507768 (AGK),
AST-0529734 (TA),
AST-0607675 (AGK),
AST-0702923 (EJH),
AST-0707474 (BDS), 
AST-0708960 (MLN), 
AST-0807075 (TA),
AST-0807215 (JB),
AST-0808184 (MLN, AGK),
AST-0808398 (TA),
AST-0908740 (AGK, DCC),
AST-0908819 (BWO), 
AST-0955300 (NJG),
AST-1008134 (GB),
AST-1109570 (AGK), 
AST-1009802 (JSO), 
AST-1102943 (MLN),
AST-1106437 (JB),
AST-1210890 (GB),
AST-1211626 (JHW),
OCI-0832662 (BWO, MLN),
OCI-0941373 (BWO),
PHY-1104819 (MLN, JB),
the CI TraCS fellowship (OCI-1048505; MJT),
and the Graduate Research Fellowship program (NJG; SWS).

### 2013-10-22 numerical-gravity.tex
Removed commented-out text from lots of files - basically vestigial junk.

% EJT: small changes only.
% dark matter particles -> massive (dark matter and star) particles

% Changed:  ``Particles from subgrids are also added in this way'' to ``Particles on subgrids within the grid's volume are also added to its gravitating field using the same method.'' I expanded because I missed the point and had to clarify with Brian when I read the text. But if it's too long, please just cut it!

% removed blank reference [REF]

@@

To compute the potential field from this gravitating mass field on the root grid, we use a a fast Fourier transform. For periodic boundary conditions, we can use either a simple Greens function kernel of $-k^{-2}$, or the finite difference equivalent \citep{Hockney88}:
\begin{equation}
G(\myvec{k}) = - \frac{\Delta x}{2 \left( \sin(k_x \Delta x/2)^2 + \sin(k_y \Delta y/2)^2 + \sin(k_z \Delta z/2)^2 \right) }
\end{equation}
where $k^2 = k_x^2 + k_y^2 + k_z^2$ is the wavenumber in Fourier space and the potential is calculated in k-space as usual with $\tilde{\phi}(k) = G(k) \tilde{\rho}(k)$.

>>>>

@@

To compute the potential field from this gravitating mass field on the
root grid, we use a a fast Fourier transform. For periodic boundary
conditions, we can use either a simple Greens function kernel of
$-k^{-2}$, or the finite difference equivalent \citep{Hockney88}:

\begin{equation}
G(\myvec{k}) = - \frac{\Delta x}{2 \left( \sin(k_x \Delta x/2)^2 + \sin(k_y \Delta y/2)^2 + \sin(k_z \Delta z/2)^2 \right) }
\end{equation}

where $k^2 = k_x^2 + k_y^2 + k_z^2$ is the wavenumber in Fourier space
and the potential is calculated in k-space as usual with
$\tilde{\phi}(k) = G(k) \tilde{\rho}(k)$.

### 2013-10-22 numerical-mhd-ct.tex
Removed commented-out text from lots of files - basically vestigial junk.

The divergence of the magnetic field, \Divb, is
identically zero in reality due to the fact that the evolution of the magnetic field is the
curl of a vector, and the divergence of the curl of a vector is identically
zero.
%\begin{eqnarray}
%\partial \Bvec/\partial t & =& \frac{1}{a} \curl \Evec\\
%\Divb & =& 0.
%\end{eqnarray}
The Constrained Transport (CT) method \citep{Evans88, Balsara99} for
magnetohydrodynamic (MHD) evolution employs this same vector property
to evolve the magnetic field in a manner that preserves $\Divb=0$.  
The electric field is computed using
the fluxes from the Riemann solver.  The curl of that electric field
is then used to update the magnetic field.  The advantage of this
method is that it preserves \Divb\ to machine precision.  The primary
drawback is increased algorithm complexity.  Note also that since 
only the update of the magnetic field is divergence free, any monopoles 
created by other numerical sources (such as ill-chosen initial conditions) persist.

@@

%Cosmological MHD can be treated in a number of ways.  As described in 
%the previous section, and equation~\ref{eq:induction} the MUSCL with Dedner
%method treats the cosmological dilution of
%$\vecB$ with a source term,  However, one can eliminate the right hand side of equation 
%\ref{eq:induction} by defining
%$\vecB^\prime = \vecB / \sqrt{a}.$
%Then equation~\ref{eq:induction} becomes
%$$ 
%\frac{\partial \vecB^\prime}{\partial t} - \frac{1}{a}  \curl (\vecv \times
%\vecB^\prime)  = 0, \label{eq:induction_cosmo_prime_nosource}
%$$
%and we can reuse each part of the non-cosmological algorithm (hyperbolic solver,
%electric field creation, interpolation, projection, and flux correction) without
%modification.  
%GB: I removed the above because the rest of the paper now uses the definition
% of B in MHD-CT.

@@

%The additional equations are
%\begin{eqnarray}
%\end{eqnarray}
%\begin{eqnarray}
%  \frac{\partial S}{\partial t} +\frac{1}{a} \div (S \vecv) & = &  -
%  \frac{ 3 (\gamma - 1) \dot{a}}{a} S \\
%  \frac{\partial \rho e}{\partial t} + \frac{1}{a} \div (\rho e \vecv) &
%  = & -\frac{ 3(\gamma -1 ) \dot{a}}{a} + \frac{p}{a} \div \vecv,
%\end{eqnarray}
%where $e$ is the internal energy and $S=p/\rho^{\gamma-1}$ is the entropy.  If
%the ratio of thermal to kinetic energy, $\eta$, is below a certain value, the
%alternative energy is used.  As in \citet{Li08a}, we use $\eta=0.008$.

%
%\enzo's MHD-CT has been used in a variety of contexts, from galaxy
%cluster evolution \citep{Xu11,Xu12}, synthetic observations of radio
%halos \citep{Skillman13}, primordial star formation \citep{Xu08}, and
%present-day star formation \citep{Collins11, Collins12a}.
% GB NOTE: incorporate these elsewhere

>>>>

The divergence of the magnetic field, \Divb, is identically zero in
reality due to the fact that the evolution of the magnetic field is
the curl of a vector, and the divergence of the curl of a vector is
identically zero.  The Constrained Transport (CT) method
\citep{Evans88, Balsara99} for magnetohydrodynamic (MHD) evolution
employs this same vector property to evolve the magnetic field in a
manner that preserves $\Divb=0$.  The electric field is computed using
the fluxes from the Riemann solver.  The curl of that electric field
is then used to update the magnetic field.  The advantage of this
method is that it preserves \Divb\ to machine precision.  The primary
drawback is increased algorithm complexity.  Note also that since only
the update of the magnetic field is divergence free, any monopoles
created by other numerical sources (such as ill-chosen initial
conditions) persist.

@@

@@

### 2013-10-22 numerical-radiation-fld.tex
Removed commented-out text from lots of files - basically vestigial junk.

%% With regards to the numerical method section, we're really looking for
%% something that's no more than a page long: a few relevant equations,
%% maybe a finite difference equation or two if you feel that it's
%% relevant, but primarily a description of the algorithms used for the
%% FLD solve and a brief explanation of how it's coupled to the
%% chemistry, as well as references to the relevant method paper(s).
%% Working in a reference or two to science papers is also great!

In addition to the ray-tracing approach for radiation transport
described in Section~\ref{sec.num.raytracing}, \enzo\ currently includes
a field-based radiation transport solver for problems posed on uniform
(i.e.~non-AMR, non-static mesh refinement) grids, which has been tuned for large-scale simulations
involving many ionizing sources.  Detailed explanations of the model and
solution approach may be found in
\citet{NBHBROW2007}, \citet{ReynoldsHayesPaschosNorman2009},
\citet{NRS2009}, and \citet{2013arXiv1306.0645N}, the salient features of which are reproduced here. 
In addition, comparisons of this solver with other astrophysical
radiation transport solvers may be found in \cite{IlievEtAl2009}.  
\enzo's field-based radiation solver focuses on
a flux-limited diffusion approximation for cosmological radiative
transfer, with couplings to both the gas energy and chemical number
densities.
%\begin{eqnarray}
%  \label{eq:fld_radiation}
%  \partial_t E_r + \frac1a \div \left(E_r \myvec{v}_b\right) &=& 
%    \div\left(D\grad E_r\right) -
%    \frac{\dot{a}}{a} E_r - c\kappa E_r + \eta, \\
%  \label{eq:fld_heating}
%  \partial_t e_c &=& -\frac{2\dot{a}}{a} e_c + G - \Lambda, \\
%  \label{eq:fld_chemistry}
%  \partial_t {\mathrm n}_i + 
%    \frac1a \div\left({\mathrm n}_i \myvec{v}_b\right) &=& 
%    \alpha_{i,j} {\mathrm n}_e {\mathrm n}_j - {\mathrm n}_i
%    \Gamma_i^{ph}, \quad i=1,\ldots,N_s,
%\end{eqnarray}
%where $E_r$ is a grey radiation energy density, $e_c$ is the
%internal energy correction due to photo-heating and chemical cooling,
%and ${\mathrm n}_i$ correspond to the chemical number densities of
%HI, HII, HeI, HeII and HeIII, and ${\mathrm n}_e$ is the electron
%number density.  Here, we define the grey radiation energy density
%through first assuming a fixed frequency spectrum, i.e.
%$E_{\nu}(\nu,\myvec{x},t) = \tilde{E}_r(\myvec{x},t) \chi(\nu)$, and then
%defining the integrated quantity
%\begin{equation}
%\label{eq:grey_radiation_energy}
%   E_r(\myvec{x},t) \equiv \int_{\nu_0}^{\infty}
%   E_{\nu}(\nu,\myvec{x},t)\,\mathrm d\nu \  = \ 
%   \tilde{E}_r(\myvec{x},t) \int_{\nu_0}^{\infty} \chi(\nu)\,\mathrm d\nu.
%\end{equation}
%In addition, $D$ in (\ref{eq:fld_radiation}) is the {\em Larsen
%square-root flux-limiter} (see \cite{Morel2000}), $\kappa$ is the total
%opacity, $\eta$ is the field of radiation sources, $G$ provides the
%radiation induced photo-heating, $\Lambda$ provides the chemical
%cooling, $\alpha_{i,j}$ are the reaction rate coefficients defining
%the interactions between species ${\mathrm n}_i$ and ${\mathrm n}_j$,
%and $\Gamma_i^{ph}$ are the radiation induced photo-ionization rates.

>>>>

In addition to the ray-tracing approach for radiation transport
described in Section~\ref{sec.num.raytracing}, \enzo\ currently
includes a field-based radiation transport solver for problems posed
on uniform (i.e.~non-AMR, non-static mesh refinement) grids, which has
been tuned for large-scale simulations involving many ionizing
sources.  Detailed explanations of the model and solution approach may
be found in \citet{NBHBROW2007},
\citet{ReynoldsHayesPaschosNorman2009}, \citet{NRS2009}, and
\citet{2013arXiv1306.0645N}, the salient features of which are
reproduced here.  In addition, comparisons of this solver with other
astrophysical radiation transport solvers may be found in
\cite{IlievEtAl2009}.  \enzo's field-based radiation solver focuses on
a flux-limited diffusion approximation for cosmological radiative
transfer, with couplings to both the gas energy and chemical number
densities.

### 2013-10-22 numerical-zeus.tex
Removed commented-out text from lots of files - basically vestigial junk.

We next examine the transport step, which is conservative.  Once again, we dimensionally split the equations and present only the one-dimensional version.  The finite difference equations actually solved are:
\begin{equation}
\rho_j^{n+d} = \rho_j^{n} - \frac{\Delta t}{\Delta x} (v^{n+c}_{j+1/2} \rho^{*}_{j+1/2} - v^{n+c}_{j-1/2} \rho^{*}_{j-1/2} )
\end{equation}
Here $\rho^*_j$ is the correctly upwinded value of $\rho$ evaluated at the cell-face corresponding to $v_j$, making $\rho^*_j v_j$ the mass flux at the cell boundary and guaranteeing mass conservation.   This requires interpolating each cell-centered quantity to the cell edge.  As recommended in \citet{Stone92a}, we use the second-order van Leer scheme, which uses piecewise linear functions.  These are given by Equations (48) and (49) of \citet{Stone92a}.  The transport steps for the other variables are similar.  Note that we advect the specific energy and specific momenta using the mass flux, as dictated by the principle of consistent transport.  This requires appropriate averaging for the momenta in the perpendicular directions as outlined in equations (57)-(72) of \citet{Stone92a}.

%GB: the following is not correct (it is based on the Anninos & Norman paper which uses a
%GB: different set of definitions for the comoving coordinates).
%\begin{eqnarray}
%Q_{ii} & = & \left\{ 
%   \begin{array}{ll}
%      Q_{\rm AV} \rho_b  (a \Delta v_{i} + \dot{a} \Delta x_i)^2 \quad
%       & \textrm{for $a \Delta v_{i} + \dot{a} \Delta x_i < 0$}  \\
%      0 & \textrm{otherwise}\\
%\end{array} \right. \\
%%\end{eqnarray}
%%and
%%\begin{equation}
%Q_{ij} & = & 0  \;\;\;{\rm for}\;\; i \ne j . 
%%\end{equation}
%\end{eqnarray}
%$\Delta x_i$ and $\Delta v_{i}$ refer to the comoving width of the grid cell
%along the $i$-th axis and the corresponding difference in gas
%peculiar velocities across the grid cell, respectively, and $a$ is the
%cosmological scale factor.

>>>>

We next examine the transport step, which is conservative.  Once
again, we dimensionally split the equations and present only the
one-dimensional version.  The finite difference equations actually
solved are:

\begin{equation}
\rho_j^{n+d} = \rho_j^{n} - \frac{\Delta t}{\Delta x} (v^{n+c}_{j+1/2} \rho^{*}_{j+1/2} - v^{n+c}_{j-1/2} \rho^{*}_{j-1/2} )
\end{equation}

Here $\rho^*_j$ is the correctly upwinded value of $\rho$ evaluated at
the cell-face corresponding to $v_j$, making $\rho^*_j v_j$ the mass
flux at the cell boundary and guaranteeing mass conservation.   This
requires interpolating each cell-centered quantity to the cell edge.
As recommended in \citet{Stone92a}, we use the second-order van Leer
scheme, which uses piecewise linear functions.  These are given by
Equations (48) and (49) of \citet{Stone92a}.  The transport steps for
the other variables are similar.  Note that we advect the specific
energy and specific momenta using the mass flux, as dictated by the
principle of consistent transport.  This requires appropriate
averaging for the momenta in the perpendicular directions as outlined
in equations (57)-(72) of \citet{Stone92a}.

### 2013-10-22 parallel.tex

To make progress, we assume that the computational cost to advance a
single cell by one timestep is a constant.  This can be incorrect if
the chemistry solver requires many iterations, but is usually fairly
accurate.  For a unigrid calculation, the time would by proportional
to $N_{\rm root}^4$ since the number of cells scales as $N_{\rm
root}^3$ and, assuming the Courant condition is the 
factor that controls the timestep, the number of steps to advance the
calculation over a given time interval is
proportional to $N_{\rm root}.$ Therefore, to advance a hierarchy a
given time interval, we find, accounting for all levels and using a refinement
factor of 2,
\begin{equation} t_{\rm SU} = C_1 \sum_l f_l N_{\rm root}^4 2^{4l}
\end{equation} where $C_1$ is a constant, which can be thought of as
the time taken to advance a single cell.  The factor $f_l$ is the
fraction of the volume on a given level that is actually refined.  By
definition $f_0 = 1$, and $f_l \le 1$.  In writing down this equation,
we neglect a number of costs that are not directly proportional to
cell count, including communication between processors, the $\log{N}$
factor for the root grid FFT, cache misses, optimizations, and other
costs associated with processing the hierarchy.  The first item on
this list, in particular, is clearly important for large processor
counts; however, we neglect parallel considerations in this section.

@@

The memory used by each level, on the other hand, scales as
\begin{equation}
{\rm memory} = C_2 \sum_l f_l N_{\rm root}^3 2^{3l}
\end{equation}
and the terms in this sum are shown in the bottom-right panel of
Figure~\ref{fig:scaling}.  This demonstrates that the memory usage is
dominated by the top three levels, and adding additional levels 
adds minimally to the memory usage of the run.  We have empirically
tested this scaling for cosmological simulations and found it to be
reasonably accurate (although the increase is typically slightly
larger than found here, usually 30-50\%, probably because of
less-than-ideal parallel scaling).

@@

Putting this together, we find an approximate scaling for the
computational time required for cosmological simulations:
\begin{equation}
t_{\rm SU} \propto M_{\rm res}^{-1} l_{\rm max}.
\end{equation}
This is approximately accurate, and gives users some way to estimate
compute times for \enzo\ calculations.

>>>>

To make progress, we assume that the computational cost to advance a
single cell by one timestep is a constant.  This can be incorrect if
the chemistry solver requires many iterations, but is usually fairly
accurate.  For a unigrid calculation, the time would by proportional
to $N_{\rm root}^4$ since the number of cells scales as $N_{\rm
root}^3$ and, assuming the Courant condition is the factor that
controls the timestep, the number of steps to advance the calculation
over a given time interval is proportional to $N_{\rm root}.$
Therefore, to advance a hierarchy a given time interval, we find,
accounting for all levels and using a refinement factor of 2,

\begin{equation} 
t_{\rm SU} = C_1 \sum_l f_l N_{\rm root}^4 2^{4l}
\end{equation}

where $C_1$ is a constant, which can be thought of as the time taken
to advance a single cell.  The factor $f_l$ is the fraction of the
volume on a given level that is actually refined.  By definition $f_0
= 1$, and $f_l \le 1$.  In writing down this equation, we neglect a
number of costs that are not directly proportional to cell count,
including communication between processors, the $\log{N}$ factor for
the root grid FFT, cache misses, optimizations, and other costs
associated with processing the hierarchy.  The first item on this
list, in particular, is clearly important for large processor counts;
however, we neglect parallel considerations in this section.

@@

The memory used by each level, on the other hand, scales as

\begin{equation}
{\rm memory} = C_2 \sum_l f_l N_{\rm root}^3 2^{3l}
\end{equation}

and the terms in this sum are shown in the bottom-right panel of
Figure~\ref{fig:scaling}.  This demonstrates that the memory usage is
dominated by the top three levels, and adding additional levels 
adds minimally to the memory usage of the run.  We have empirically
tested this scaling for cosmological simulations and found it to be
reasonably accurate (although the increase is typically slightly
larger than found here, usually 30-50\%, probably because of
less-than-ideal parallel scaling).

@@

Putting this together, we find an approximate scaling for the
computational time required for cosmological simulations:

\begin{equation}
t_{\rm SU} \propto M_{\rm res}^{-1} l_{\rm max}.
\end{equation}

This is approximately accurate, and gives users some way to estimate
compute times for \enzo\ calculations.
